\chapter*{Introducing {\Huge VAMPIRE}}
\addcontentsline{toc}{chapter}{Preface}
\phantomsection\addcontentsline{toc}{section}{Introducing VAMPIRE}

\vampire is a state-of-the-art atomistic simulator for magnetic nanomaterials.
This software is the culmination of several years of continuous development,
with an aim to make atomistic simulation of magnetic materials routinely
available to the non-specialist researcher. Before now, using atomistic
models to simulate magnetic systems required in depth and technical knowledge
of the underlying theoretical methods, computer programming skills and the
ability to debug and understand intricate computational problems. The code is
designed with ease of use in mind, and includes an extensive set of input
parameters to control the simulations through a plain text input file. Subject
to future funding it is also hoped to develop graphical user interfaces for
macOS$^{\mathrm{TM}}$ and Windows$^{\mathrm{TM}}$ which should make using
the code more accessible.

The \vampire project is still very much under active development, with an open
development of all code features. The features are always available during the
development stages from the develop branch of the code, but with the caveat that
they are not always fully reliable. Feedback of any bugs or errors to the
\vampire developers is always welcome, as well as any feature requests or
enhancements.

We hope that as the \vampire project develops it will become a useful tool for
the magnetics community for specialists and non-specialists alike.